\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the PRM Method

### 2015-12-24 description/description.tex
prm projekt angefangen

\section{Problem Statement and Sequential PRM Method}
\newtheorem*{defi}{Definition}
\newtheorem*{alg1}{Algorithm}

@@

The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first select randomly a node \(q_0 \in V\) and then sample a new node \(q\) in the neighbourhood of \(q_0\). Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.

@@

In the project, it is planned to use this algorithm to create collisionfree trajectories of a robot arm. 
For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot. Therefore the direct kinematic can be interpreted as a (strongly nonlinear) function
\begin{equation*}
	f: (q_1, ... , q_d) \mapsto (x_1, R_1, ..., x_l, R_l).
\end{equation*}

Now, we have given some obstacles with fixed positions \(x_{l+1}, ..., x_m\) and orientations \(R_{l+1}, ..., R_{m}\), which the robot has to avoid during movement. It is assumed that all objects are given in a form, that for each two objects it can be determined from the positions and orientations, if there is an intersection. For example as polyhedrons.
From this the indicator function can be defined as
\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{if no collision for all object pairs\ } (x_j, R_j), (x_k,R_k),\ j,k\in 1, ... , m \\
		1,	& \text{for at least one collision}
	\end{cases}
\end{equation*}
Now, if we additionally introduce jointlimits and call the set of possible angles \(\Omega\), we see, that the problem of finding feasible trajectories is exactly the pathfinding problem from above.

@@

Like the stated robotics example, in many cases the most costly operation of the PRM algorithm is the evaluation of the indicator function \(I(q)\), which however can be done independently for every new \(q\).
\begin{itemize}
\item 
Therefore in a first step, it is planned to implement \(I(q)\) parallely in a vectorized way, i.e. as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\). For example with help of the GPU. In the robotics example this means, the direct kinematic and collision calculation is made parallely for \(P\) different joint angle values.
\item
Then in a first level of parallelization, this function can be applied to the \(q_i \in [q, \tilde q]_h \) to check, if \(q\) and \(\tilde q\) are connected.
\item
Furthermore, a second level of parallelization could be achieved by sampling and connecting several new points \(\tilde q_i\) parallely. This is expected to be more difficult, because all new nodes have to be connected to the old graph and to each other. Therefore it has to be evaluated, if a parallel implementation stays efficient here.
\item
If there remains time, a next idea could be to execute the algorithm on different discretization levels \(h\). This means for example to compute a way \(q_1^0, ..., q_n^0\) which is connected with a coarse stepsize \(h_0\), then to decrease the stepsize to \(0<h_1<h_0\) and take the path nodes \(q_i^0\) as starting nodes for the next execution of the algorithm, which gives then an \(h_1\)-connected way \(q_1^1, ... , q_n^1\), etc.
\end{itemize}

>>>>

\section{Problem Statement and Sequential PRM Method}\label{algintro}
\newtheorem*{defi}{Definition}
\newtheorem*{alg1}{Algorithm}

@@

The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\). Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.

@@

In the project, it is planned to use this algorithm to create collisionfree trajectories of a robot arm. 
For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot, \cite{robodyn} describes the exact calculations. Therefore this direct kinematic can be interpreted as a (strongly nonlinear) function
\begin{equation*}
	f: (q_1, ... , q_d) \mapsto (x_1, R_1, ..., x_l, R_l).
\end{equation*}

Now, we have given some obstacles with fixed positions \(x_{l+1}, ..., x_m\) and orientations \(R_{l+1}, ..., R_{m}\), which the robot has to avoid during movement. It is assumed that all objects are given in a form, that for each two objects it can be determined from the positions and orientations, if there is an intersection. For example as polyhedrons.
From this the indicator function for the obstacles in the configuration space can be defined as
\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{if no collision for all object pairs\ } (x_j, R_j), (x_k,R_k),\ j,k\in 1, ... , m \\
		1,	& \text{for at least one collision}
	\end{cases}
\end{equation*}
Now, if we additionally introduce jointlimits and call the set of possible angles \(\Omega\), we see, that the problem of finding feasible trajectories is exactly the pathfinding problem from above.

@@

Like in the stated robotics example, in many cases the most costly operation of the PRM algorithm is the evaluation of the indicator function \(I(q)\), which however can be done independently for every new \(q\).
\begin{itemize}
\item 
Therefore in a first step, it is planned to implement \(I(q)\) parallely in a vectorized way, i.e. as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\). For example with help of the GPU. In the robotics example this means, the direct kinematic and collision calculation is made parallely for \(P\) different joint angle vectors.
\item
Then in a first level of parallelization, this function can be applied to the \(q_i \in [q, \tilde q]_h \) to check, if \(q\) and \(\tilde q\) are connected.
\item
Furthermore, a second level of parallelization could be achieved by sampling and connecting several new points \(\tilde q_i\) parallely. This is expected to be more difficult, because all new nodes have to be connected to the old graph and to each other. Therefore it has to be evaluated, if a parallel implementation stays efficient here.
\item
If there remains time, a next idea could be to execute the algorithm on different discretization levels \(h\). This means for example to compute a way \(q_1^0, ..., q_n^0\) which is connected with a coarse stepsize \(h_0\), then to decrease the stepsize to \(0<h_1<h_0\) and take the path nodes \(q_i^0\) as starting nodes for the next execution of the algorithm, which gives then an \(h_1\)-connected way \(q_1^1, ... , q_n^1\), etc.
\end{itemize}
